### 2013-11-25 paper/paper.tex
Beginning of Dan's paper revisions

\documentclass[12pt]{article} 
\usepackage{graphics}
\usepackage{setspace}
\usepackage{cite}

% the following is to get inch margins on a letter-size paper
\setlength{\topmargin}{0pt}
\setlength{\headheight}{0in}
\setlength{\headsep}{0in}
\setlength{\textheight}{9.0in}
\setlength{\footskip}{0.5in}
\setlength{\oddsidemargin}{0pt}
\setlength{\evensidemargin}{0pt}
\setlength{\textwidth}{6.5in}

\doublespacing
\begin{document}
\title{A Java Wrapper For Embarrassingly Parallel Problems}
\author{Jacob Schwartz}
\maketitle

\begin{abstract}
An embarrassingly parallel problem, also called pleasingly parallel, is one that
can be easily broken up into identical components that do not need to interact
to find the solution. Several problems in the biological sciences are
embarrassingly parallel but they are difficult to rewrite to use multithreading,
because they are either poorly written or too complex. We have written a
framework, implemented in Java, that will execute the serial program in several
threads. These threads working together help cut down the total runtime of the
program exponentially. The next step is to separate the framework into a
front-end that will run locally and a back-end which receives blocks of work and
does the computations. This back-end should run in the cloud, on a machine that
is dedicated to running embarrassingly parallel tasks only.
\end{abstract}

\section{Introduction}

A computation graph consists of nodes that represent the functions or processes.
The edges of the graph show data being passed from one process to another. If
the graph is fully disconnected, none of the processes need to communicate to
complete the computations. Such a computation is called embarrassingly
parallel.\footnote{The term embarrassingly parallel was coined by Cleve Moler in
his 1986 paper ``Matrix Computation on Distributed Memory Multiprocessors.''} In
a practical situation, there will be some inter-node communication to set up the
problem, and then again to accumulate the results.  Embarrassingly parallel
problems can be easily performed on a cluster of servers or in any kind of
distributed network, due to its lack of dependencies.

Programs in the field of Biology and Genetics often analyze long strings of
characters, whether they are DNA, RNA or protein sequences. In addition to
handling large file sizes, there is a lot of processing that goes into comparing
and manipulating strings. This time adds up when there are hundreds of thousands
of strings to process. However, these sequences often do not relate to each
other and can be processed independently. Using an embarrassingly parallel
approach, the processing time for all of the sequences can often be cut down
exponentially with very minimal effort by the programmer.

This paper describes a framework that makes it easy to parallelize a potentially
embarrassingly parallel program that reads from a file, writes to a file and
whose input can be broken up into work units in an easily identifiable manner.
The framework provides a ``wrapper'' for the original program that mimics the
behavior of the original program. It calls the serial programs on multiple Java
threads, such that each thread acts like the serial application, reading a
subset of the input in the same manner and creating the output in the same
format. The framework can result in a significant exponential decrease with very
little programmer effort.

\section{Background}

TODO: Arrows?
The motivation for this effort was based on research into different forms of the
species \emph{Cicer arietinum}, the chickpea. The two main forms, Kabuli and
Desi, are very different in size and color. The Desi is also more resistant to
diseases and are more effective at root nodule symbiosis. The Kabuli has a large
but soft shell, which make it at better food source for humans. Both forms are
cultivated and have been under human selection for several thousand years. We
wanted to use an existing Ka/Ks program (Zhang et al. \cite{kaks}) to quantify
the evolution and selection of these chickpeas. We needed to analyze three
variants, each with thousands of sequences. The Ka/Ks Calculator program took
between five to thirty seconds to process one pair of sequences on a commodity
desktop.

At this rate, we predicted it would take ten days of constant computation to
complete the processing we needed. We neither had the time nor the expertise to
develop a new parallel Ka/Ks program. We were also reluctant to revise the
existing code to to implement the parallel computation because of the danger of
introducing bugs into unfamiliar code and because of the learning curve required
for C++ threading. Instead we wrote, a simple Java program to break up the input
and spawn threads that would run the Ka/Ks Calculator process without having to
make any changes to the code. When al of the threads ended, the contents of the
output files were concatenated together to produce a final output that matches
the format and order that would have been produced by a single invocation of the
Ka/Ks Calculator program.

%TODO: get a better source? Talk about UCLUST and octupus
Ka/Ks Calculator is not the only embarrassingly parallel program that has the
potential to take several days to run on a data set. In fact, there are several
embarrassingly parallel programs in the biological sciences. To deal with large
amounts of data, the best method is often to group them into clusters.
Hierarchical clustering algorithms take the data and then organize into
hierarchical structure, usually a dendrogram, based on a priority matrix. There
are hierarchical clustering algorithms that can be implemented in parallel using
an embarrassingly parallel approach.  With several applications that would
benefit from parallelism in this field, it was obvious it would be desirable to
develop a framework for the  parallelization task. \cite{cluster}

%TODO: TALK ABOUT CLOUD PAPER HERE.
Segue. Gunarathne et al.\cite{cloud} have investigated running much larger
embarrassingly parallel biological problems using various cloud services. These
services offer on-demand computational power with MORE.
These services also have built in hooks to various distributed services, such as
storage and databases. One of these technologies is the MapReduce framework,
which allows user to perform distributed computation similar to this framework.
As seen in figure X, Apache Hadoop, an open-source implementation of MapReduce,
breaks down the input data into parts, runs them in parallel with the 'Map'
function and then combines the results with the 'Reduce' function. TWO MORE
SENTENCES.

\section{Proposed Solution}
\begin{figure}
{\resizebox{5.2in}{!}{\includegraphics{figures/uml.png}}}
\caption{UML for the embarrassingly parallel framework}
\label{fig:uml}
\end{figure}
The only requirements for a program to be wrapped by framework are that it 
process the input in an embarrassingly parallel manner and that the input comes 
from a file. The processed results must also written to a file. The framework 
supports two forms of input: a directory of files that contain that each contain
one unit of work, which is represented by the Chunk class, or a single file that
the framework can break up into Chunks. For example, the Ka/Ks Calculator's 
input consists of a header and then two protein sequence lines followed by a 
blank line. The framework uses that blank line to know when one Chunk of work 
has ended and another has begun. The user can input a regular expression to 
define when lines of input are separated into Chunks. This occurs in the 
ChunkManager class.

The other purpose of the ChunkManager class is to combine the resulting Chunks
back together when the worker threads have finished. Once the work has been
done, all of the separate results are spread across several files. Those results
must be retrieved and combined to construct the final output file. This output
file must look the same as the output file would look if the program were run
serially; the output may need to come in the same order as the inputs in the
input file and headers may need to be deleted from the separate files and added
to the combined file. End-users can also choose to write their own merge code in
Java, using the CustomMerge class, or write their own independent merge script.

The end-user can control many aspects of the parallelization of the code,
including the number of threads to use, various flags for the executable and the
statistics output.  A start-up wizard allows the end-user to choose the settings
for the run and creates an instance of Configuration when it is completed to
save the user's settings, allowing the user to rerun with the same settings
without having to go through the wizard again.

Lastly, the worker threads keep records during execution. The threads record
their total uptime and the number of Chunks they execute, and the runtime for
each piece of work. This information will be used to evaluate the effectiveness
of the approach, but can also provide useful information the can help a user
tine a particular application on a particular type of input data. The statistics
for each thread can be found in the Worker subclass and the individual Chunk
objects will store their own statistics about runtime and size.

\section{Results}

We evaluated the convenience and performace of the framework on two different
applications: the Ka/Ks Calculator and two different sequence clustering
algorithms. In all cases, we compared the framework based application with the
original using different data partitions and different thread counts.

\subsection{Ka/Ks Calculator}
Ka/Ks Calculator was run through the framework with ten different samples, each 
containing sixteen units of work. Each sample was run by the framework thread
counts, starting at 1 and increasing by 
powers of two until the 100\% processor usage (16 processors). These tests were run
with minimal other activity on the server.

\begin{figure}
{\resizebox{5.2in}{!}{\includegraphics{experiment/out.png}}}
\caption{Runtimes for the Ka/Ks Calculator though the framework}
\label{fig:graph}
\end{figure}

Figure 2 summarized the results; each line shows the time taken for each chunk
of the data.
As the numbers of threads increases, the runtime of the program decreases 
exponentially. In a few cases, there was an increase in time when attempting to 
max out the CPU usage. This occurs because the framework has to wait for the 
CPU to be finished with its previous work before it can start any 
further jobs.

TODO(Keep?): In the cases when the framework executed faster while the CPU was 
maxed out, it was only faster a minute at the most. However, running the 
framework at 50\% of the CPU is more efficient than running using 100\% of the 
CPU. The runtime drop achieved when using 100\% of the CPU versus 50\% is not 
large enough to warrant using all of the machine's power.

\subsection{uClust}

Place holder text

\subsection{Octupus}

Place holder text

\section{Conclusion}

TODO: First paragrpah should summarize performce: its clear that framework
provices simple and effective module for adding parallelism to a class of
file-oriented EP apps. More self praise.

Using Java was a good choice for implementation ease and future implementation
extension by others, but it may not produce the fastest results. Languages like
Hadoop, Clojure or other multithread oriented language may have been a better
choice due to their inherit parallel nature. In addition to a single machine
working to process the data quickly, we can also look at distributed solutions
by using MPI or a cloud solution like Amazon instances. The MapReduce framework
works similarly to our framework, there is a step that splits the data, the Map
function is the user's executable and the merge step is the Reduce function in
the framework.

In the future, we would like to see this work be split into a
local front-end and a cloud back-end. This allows for an even greater runtime
decrease because the number of compute nodes available is larger than the number
of processors on most machines. Splitting the framework also allows for better
testing of cloud environments, where the cloud environment itself could be a
configurable option for end users.
TODO: MORE

\begin{thebibliography}{1}
\bibitem{history}
Moler, Cleve. ``\emph{Matrix Computation on Distributed Memory Multiprocessors."
}, Hypercube Multiprocessors, 1986: Proceedings of the First Conference on
Hypercube Multiprocessors, Knoxville, Tennessee, August 24-27, 1985. By Michael
T. Heath. Philadelphia: SIAM, 1986. N. pag. Print.
\bibitem{kaks}
Zhang Z, Li J, Zhao XQ, Wang J, Wong GK, Yu J., \emph{Ka/Ks Calculator: 
calculating Ka and Ks through model selection and model averaging},
Genomics Proteomics Bioinformatics, November 2006.
\bibitem{cluster}
Rua, Xu, Wunsch, Donald II, \emph{Survey of Clustering Algorithms},
IEEE Transactions on Neural Networks VOL. 16, NO. 3, MAY 2005
\bibitem{cloud}
Crap, crap and Crap
\end{thebibliography}

\end{document}